Supplementary Materials

### 2022-12-09 manuscript/supplementary.tex
ms supp updates

\subsection{Estimates of Pairwise Diversity}

Our underlying data for all likelihood and pairwise diversity estimates is the
allele count matrix $\mathbf{C}$ with dimensions $L \times 2$, where $L$ is the
chromosome length. This is converted to a pairwise summary matrix with
identical dimensions, $\mathbf{P}$. The first column of $\mathbf{P}$ is the
number of pairwise comparisons between chromosomes that are identical, and the
second column is the number that are different. Both of these columns are
combinatoric summaries of the allele counts matrix. Let $[c_1, c_2]$ be a row
of $\mathbf{C}$ for basepair $l$, and $n = c_1 + c_2$. Then, the (1) total
number of pairwise combinations of chromosomes $n_T$, (2) the number of
pairwise with identical alleles $n_S$, and (3) the number of pairwise
combinations with differing alleles $n_D$ are respectively,

\begin{align*}
  n_T &= \frac{n(n-1)}{2} \\
  n_S &= {c_1 \choose 2} + {c_2 \choose 2} \\
  n_D &= n_T - n_S
\end{align*}

Thus $\mathbf{P}_l = [n_S, n_D]$, and the pairwise diversity across the $n$
sampled chromosomes is calculated as

@@

@@

\begin{figure}[!htb]
  \centering
  \includegraphics{figures/supplementary/diversity_trimming_dist.pdf}
  \caption{}
  \label{suppfig:trimming}
\end{figure}

>>>>

\subsection{Data Summary Matrices}

Our underlying data for all likelihood and pairwise diversity estimates is the
allele count matrix $\mathbf{C}$ with dimensions $L \times 2$, where $L$ is the
chromosome length. This is transformed to a pairwise summary matrix with
identical dimensions, $\mathbf{Y}$. The first column of $\mathbf{Y}$ is the
number of pairwise comparisons between chromosomes that are identical, and the
second column is the number that are different. Both of these columns are
combinatoric summaries of the raw allele counts matrix needed for the binomial
likelihood and pairwise diversity estimates. Let $[c_1, c_2]$ be a row of
$\mathbf{C}$ for basepair $l$ (the $l$ index is omitted for clarity), and $n =
c_1 + c_2$. Then, the (1) total number of pairwise combinations of chromosomes
$n_T$, (2) the number of pairwise with identical alleles $n_S$, and (3) the
number of pairwise combinations with differing alleles $n_D$ are respectively,

\begin{align*}
  n_T &= \frac{n(n-1)}{2} \\
  n_S &= {c_1 \choose 2} + {c_2 \choose 2} \\
  n_D &= n_T - n_S
\end{align*}
%
which would be stored in row $\mathbf{Y}_l = [n_S, n_D]$.

\subsection{Pairwise Diversity Estimates}

The pairwise diversity at site $l$ across the $n$ sampled chromosomes can be
calculated from the $\mathbf{Y}$ matrix as follows,

@@

The genome-wide $\bar{\pi}$ can also be estimated by summing the columns of
$\mathbf{Y}$, and calculating $\nicefrac{n_D}{n_T}$.

@@

\begin{figure}[!htb]
  \centering
  \includegraphics{figures/supplementary/diversity_trimming_dist.pdf}

  \caption{The distribution of diversity across the genome at the megabase
  scale, with outliers trimmed. The blue line is the normal CDF, fit with MLE
parameters XXX. The inset figure is the untrimmed data, with trimmed points
shown in red.}

  \label{suppfig:trimming}
\end{figure}
